Self-interacting dark matter model without dark energy in cosmology

### 2025-03-05 article/article.tex
transfer the t0 and zmax

   However, the solution of Eq.(\ref{eq:8}) gives an upper limit of redshift $z_{\max}$
   (see Appendix), so the IDM model cannot give a constraint to the $r_{\text{d}}$ and we just try to use the cross
   parameter $r_{\text{d}}h$ to give the constraints.

@@

@@

   \begin{table*}[htbp]
      \caption{The constraint results of Model 1}
      \centering
      \begin{tabular}{lcccccc}
         \hline\hline
         Model 1 & OHD & SNe Ia & SNe Ia + QSO & BAO &
          OHD + SNe Ia & OHD + SNe Ia + BAO \\
         \hline
         $\Omega_{2,0}$ & $0.321_{-0.067}^{+0.055}$ & $0.360_{-0.020}^{+0.020}$
          & $0.359_{-0.020}^{+0.020}$ & $0.301_{-0.013}^{+0.013}$
          & $0.330_{-0.016}^{+0.016}$ & $0.3131_{-0.0099}^{+0.0099}$ \\
         $\log_{10}(\kappa C_1/$Gyr${}^{-1})$ & < & < & < & < & < & < \\
         $\log_{10}(M_x/$GeV) & > & > & > & > & > & > \\
         \hline
      \end{tabular}
      \label{tab:5}
   \end{table*}

@@

@@

   Apply the Eq.(\ref{eq:12}) to Eq.(\ref{eq:8}), we can get a simple nonlinear
   second-order differential equation for the redshift $z(t)$, which can
   be written as \begin{equation}
      2H_0^2\Omega_{1,0}(1+z)^2z'z''+\kappa C_1z'{}^4-5H_0^2\Omega_{1,0}(1+z)z'{}^3
      +3H_0^4\Omega_{1,0}^2(1+z)^3z'-H_0^4\Omega_{1,0}^2\kappa C_1[(z+1)^4-1]
      =H_0^4\Omega_{1,0}^2\kappa C_1,
   \end{equation}
   where the prime denotes the derivative with respect to $t$.
   The equation can be simplified as
   \begin{equation}
      2H_0^2\Omega_{1,0}y^2y'y''+\kappa C_1y'{}^4-5H_0^2\Omega_{1,0}yy'{}^3
      +3H_0^4\Omega_{1,0}^2y^3y'-H_0^4\Omega_{1,0}^2\kappa C_1y^4=0,
   \end{equation}
   where $y=z+1$, now it is a homogeneous function and we can give the
   general solution as $y=\exp{f}$ with $f$ a function of $t$, the equation
   now is \begin{equation}
      2H_0^2\Omega_{1,0}f'f''+\kappa C_1f'{}^4-3H_0^2\Omega_{1,0}f'{}^3
      +3H_0^4\Omega_{1,0}^2f'-H_0^4\Omega_{1,0}^2\kappa C_1=0.
   \end{equation}
   We use $g=f'$, which also gives $g(t)=y'/y=-H(t)$, then the equation can be written as
   \begin{equation}
      2H_0^2\Omega_{1,0}gg'+\kappa C_1g^4-3H_0^2\Omega_{1,0}g^3
      +3H_0^4\Omega_{1,0}^2g-H_0^4\Omega_{1,0}^2\kappa C_1=0.
   \end{equation}
   It can be theoretically solved and the solution is like
   \begin{equation}
      g(t)=\mathcal{G}\left(-\frac{t}{2H_0^2\Omega_{1,0}}+\mathrm{Const}\right),
   \end{equation}
   where $\mathcal{G}$ is the inverse function of
   \begin{equation}
      \mathcal{F}(x)=\frac{3H_0\sqrt{\Omega_{1,0}}\mathrm{arctanh}\frac{x}{H_0\sqrt{\Omega_{1,0}}}
      -\kappa C_1\log(x^2-H_0^2\Omega_{1,0})+\kappa C_1\ln(\kappa C_1x^2-3H_0^2\Omega_{1,0}x
      +H_0^2\Omega_{1,0}\kappa C_1)}{9H_0^4\Omega_{1,0}^2-4H_0^2\Omega_{1,0}\kappa C_1^2},
   \end{equation}
   in this term, $\mathrm{arctanh}(x)\doteq \frac{1}{2}\log[(x+1)/(x-1)]$ when $x<-1$,
   and the Const is determined by the initial condition $g(t_0)=-H_0$.

   The function has a limit when $x\to-\infty$,
   \begin{equation}
      \lim_{x\to-\infty}\mathcal{F}(x)\equiv \mathcal{F}(-\infty)=
      \frac{\kappa C_1\ln\kappa C_1}{H_0^2\Omega_{1,0}(-4\kappa C_1^2+9H_0^2\Omega_{1,0})},
   \end{equation}
   which means that the IDM model can only give a solution to the lower limit $t_{\min}$,
   and the redshift $z_{\max}$ can be calculated as
   \begin{equation}
      z_{\max}=\exp\left[\int_{-\infty}^{-H_0}\left(\mathcal{F}(-\infty)
      -\mathcal{F}(x)\right)\mathrm{d}x+\frac{t_0-t_{\min}}{2H_0\Omega_{1,0}}\right]-1,
   \end{equation}
   and $z_{\max}$ is also a function of $\Omega_{2,0}, \kappa C_1$ and $H_0$.
   \begin{figure}[htbp]
      \centering
      \includegraphics[width=0.7\textwidth]{zmax.pdf}
      \caption{The redshift $z_{\max}$ as $H_0=70$ km/s/Mpc}
   \end{figure}

>>>>

   However, the solution of Eq.(\ref{eq:8}) only gives an upper limit of redshift $z_{\max}$
   (see Appendix), and we just try to use the cross
   parameter $r_{\text{d}}h$ to give the constraints.

@@

   We would find that the SNe Ia and QSO combined result is most
   similar to SNe Ia result, what may refer that the QSO cannot
   give an effective constraint.

@@

   \begin{table*}[htbp]
      \caption{The constraint results of Model 1}
      \centering
      \begin{tabular}{lcccccc}
         \hline\hline
         Model 1 & OHD & SNe Ia & SNe Ia + QSO & BAO &
          OHD + SNe Ia & OHD + SNe Ia + BAO \\
         \hline
         $\Omega_{2,0}$ & $0.321_{-0.067}^{+0.055}$ & $0.360_{-0.020}^{+0.020}$
          & $0.359_{-0.020}^{+0.020}$ & $0.301_{-0.013}^{+0.013}$
          & $0.330_{-0.016}^{+0.016}$ & $0.313_{-0.010}^{+0.010}$ \\
         $\log_{10}(\kappa C_1/$Gyr${}^{-1})$ & $<-1.81$ & $<-2.30$ & $<-2.35$
          & $<-2.66$ & $<-2.78$ & $<-3.21$ \\
         $\log_{10}(M_x/$GeV) & $>-10.09$ & $>-9.62$ & $>-9.60$
          & $>-9.25$ & $>-9.15$ & $>-8.70$ \\
         \hline
      \end{tabular}
      \label{tab:5}
   \end{table*}

@@

   The results from three dataset are different

@@

   Apply the Eq.(\ref{eq:12}) to Eq.(\ref{eq:8}), we can get a simple nonlinear
   second-order differential equation for the redshift $z(t)$, which can
   be written as \begin{equation}
      2H_0^2\Omega_{1,0}(1+z)^2z'z''+\kappa C_1z'{}^4-5H_0^2\Omega_{1,0}(1+z)z'{}^3
      +3H_0^4\Omega_{1,0}^2(1+z)^3z'-H_0^4\Omega_{1,0}^2\kappa C_1[(z+1)^4-1]
      =H_0^4\Omega_{1,0}^2\kappa C_1,
   \end{equation}
   where the prime denotes the derivative with respect to $t$.
   The equation can be simplified as
   \begin{equation}
      2H_0^2\Omega_{1,0}y^2y'y''+\kappa C_1y'{}^4-5H_0^2\Omega_{1,0}yy'{}^3
      +3H_0^4\Omega_{1,0}^2y^3y'-H_0^4\Omega_{1,0}^2\kappa C_1y^4=0,
   \end{equation}
   where $y=z+1$, now it is a homogeneous function and we can give the
   general solution as $y=\exp{f}$ with $f$ a function of $t$, the equation
   now is \begin{equation}
      2H_0^2\Omega_{1,0}f'f''+\kappa C_1f'{}^4-3H_0^2\Omega_{1,0}f'{}^3
      +3H_0^4\Omega_{1,0}^2f'-H_0^4\Omega_{1,0}^2\kappa C_1=0.
   \end{equation}
   We use $g=f'$, which also gives $g(t)=y'/y=-H(t)$, then the equation can be written as
   \begin{equation}
      2H_0^2\Omega_{1,0}gg'+\kappa C_1g^4-3H_0^2\Omega_{1,0}g^3
      +3H_0^4\Omega_{1,0}^2g-H_0^4\Omega_{1,0}^2\kappa C_1=0.
   \end{equation}
   It can be theoretically solved and the solution is like
   \begin{equation}
      g(t)=\mathcal{G}\left(-\frac{t}{2H_0^2\Omega_{1,0}}+Const\right),
   \end{equation}
   where $\mathcal{G}$ is the inverse function of
   \begin{equation}
      \mathcal{F}(x)=\frac{3H_0\sqrt{\Omega_{1,0}}\mathrm{arctanh}\frac{x}{H_0\sqrt{\Omega_{1,0}}}
      -\kappa C_1\log(x^2-H_0^2\Omega_{1,0})+\kappa C_1\log(\kappa C_1x^2-3H_0^2\Omega_{1,0}x
      +H_0^2\Omega_{1,0}\kappa C_1)}{9H_0^4\Omega_{1,0}^2-4H_0^2\Omega_{1,0}(\kappa C_1)^2}.
   \end{equation}
   In this term, $\displaystyle\mathrm{arctanh}(x)\doteq \frac{1}{2}\log\left(\frac{x+1}{x-1}\right)$
   when $x<-1$, and the $Const$ is determined by the boundary condition $g(0)\to -\infty$, which is

   \begin{equation}
      Const=\mathcal{F}(-\infty)\equiv\lim_{x\to-\infty}\mathcal{F}(x)=
      \frac{\kappa C_1\ln\kappa C_1}{H_0^2\Omega_{1,0}[-4(\kappa C_1)^2+9H_0^2\Omega_{1,0}]}.
   \end{equation}
   The initial time of today $t_0$ can be easily given as
   \begin{equation}
      t_0=2H_0^2\Omega_{1,0}[\mathcal{F}(-\infty)-\mathcal{F}(-H_0)],
   \end{equation}
   and the redshift $z_{\max}$ can be calculated as
   \begin{equation}
      z_{\max}=\exp\left[\int_{-\infty}^{-H_0}[\mathcal{F}(-\infty)
      -\mathcal{F}(x)]\mathrm{d}x+\frac{t_0}{2H_0\Omega_{1,0}}\right]-1,
   \end{equation}
   and $z_{\max}$ is also a function of $\Omega_{2,0}, \kappa C_1$ and $H_0$.
   \begin{figure}[htbp]
      \centering
      \includegraphics[width=0.7\textwidth]{zmax.pdf}
      \caption{The redshift $z_{\max}$ as $H_0=70$ km/s/Mpc}
   \end{figure}
